Deep Learning

### 2015-01-07 2014/nature-review/deepreview.tex
working on references

So far, most applications of deep learning use feed-forward neural
networks, but for tasks that involve sequential inputs it is often better
to use recurrent neural networks. These are given one element from an input % YB term-->element
sequence at each time step and their recurrent internal ``hidden'' state
depends on the whole history of the input sequence. Recurrent networks are
very good at predicting the next character in a string\cite{Sutskever} or
the next word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
more complex tasks.  After reading an English sentence one word at a time,
for example, an English``encoder'' network can be trained so that the final
state vector of its hidden neurons is a good representation of the thought
expressed by the sentence.  This thought vector can then be used as the
initial hidden state of a trained French ``decoder'' network which outputs
a probability distribution for the first word of the French translation. If
a particular first word is chosen (e.g., according to that probability)
and provided as input to the decoder
network it will then output a probability distribution for the second word
of the translation and so on until a full stop is chosen~\cite{Vinyals,
  Montreal}. 
%YB added a clarification
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

@@

Patterns can be classified by first converting each raw input vector into a
vector of feature activations and then collecting evidence for each
possible class by computing a weighted sum of the features. If good
features can be designed by hand, it is relatively easy to learn
class-specific weights that optimize
discrimination~\cite{perceptrons}. Widely-used examples of such linear
classification methods are logistic regression and linear support vector
machines~\cite{logisticregression,svm}.

However, even with good hand-crafted features, many interesting tasks
cannot be solved by linear classifiers. A common method to get around this
problem is to transform the feature vector into a higher dimensional vector
through a non-linear mapping, exploiting the fact that collections of
vectors are more likely to be easily spearated when they are spread around
a high-dimensional space~\cite{cover}. One clever way to do so is to use a
``kernel function'' that measures the similarity between two vectors.
Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input
vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
optimize the choice of training inputs used to define features and the
class-specific weights associated with the features~\cite{svm}. The training
inputs used to define features are called ``support vectors'' and Support
Vector Machines\cite{Vapnik} have been very successful, especially for
tasks with a relatively small number of training examples in which features
are individually informative. They do not, however, perform well for tasks
such as object recognition in which the input space is high-dimensional and
the input-output mapping is so complex that computing similarities with a
simple kernel function is ineffective.

@@

From the earliest days of pattern recognition~\cite{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s\cite{Werbos74,Parker85,LeCun85,Rumelhart86b,schmidreference}.  Backpropagation is simply an application of
the chain rule, which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights in a layered
network. The objective function typically measures the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it outputs} of a layer can be computed by working
backwards from the error derivatives with respect to the outputs of the
subsequent layer. Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the parameters (or weights) of each layer (see box
1).

@@

Backpropagation provides an efficient way of computing gradients for all of
the weights in a layered network on a single training case, but there are
still many ways to make use of these gradients. The most obvious method is
to average the gradients over all of the cases in the training set and then
to change each weight in proportion to its negative gradient, which is the
direction where a small change will most improve the objective function. 
When the training set contains millions of cases and is highly redundant, however,
this is needlessly slow. It is much faster to compute the average gradient
on a small ``mini-batch'' of randomly chosen training cases and to update
each weight by this stochastic estimate of its full gradient times a
learning rate which decays as learning proceeds. Stochastic gradient
descent can be improved by using the history of the individual weight
gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
or to provide additional increments that depend on earlier gradients\cite{momentum}
but these techniques are still the subject of controversy and are beyond
the scope of this review.  However, it is now widely accepted that some
form of stochastic gradient descent is the best bet for large, redundant
datasets.

@@

The connectivity of the network is restricted so that each of the one-of-N
input vectors must first be converted to a ``word-vector'' before being
used by subsequent layers to give a high probability to the target next
word.  The network learns word-vectors that contain many active components
each of which can be interpreted as a separate feature of the word. This
was first demonstrated\cite{RHW} using sequences of three words, such as
(Colin, Mother, Victoria), derived from two family trees. The learned word
vector for Colin contained ``semantic'' features representing which family
tree Colin was in, what branch of that tree he was in, and from what generation
he was~\cite{Hinton-et-al-PDP1986}. These semantic features were not explicitly present
in the input.  They were discovered by the learning procedure as a good way
of factorizing the structured relationships between the input and output
symbols into multiple ``micro-rules''.  For example, given Colin's
generation and the learned feature of Mother that the output should be one
generation higher than the input, we can already restrict the answer to
people in a specific generation. The ability of deep learning to discover
many such micro-rules allows it to generalize well to novel combinations of
individually familiar context words.

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or vanish~\cite{Bengio-et-al-TNN1994,Hochreiter91-small}.

@@

The obvious weakness of gradient descent is that it can get trapped in poor
local minima, but in high-dimensional weight-spaces this is much less
common than getting stuck close to saddle points where the error surface
curves upwards in most directions but downwards in a few. Recent
results~\cite{Dauphin-et-al-NIPS2014-small,Yann} strongly suggest
that in deep networks almost all of these local minima give very similar
performance so it does not matter which one we find.  This goes a long way
towards explaining why stochastic gradient descent works so well for deep
neural networks.

Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by
the introduction of an unsupervised learning procedure that could create
layers of feature detectors without requiring labeled data. The objective
in learning each layer of feature detectors was to be able to reconstruct
or model the activities of feature detectors (or raw inputs) in the layer
below.  By ``pre-training'' several layers of progressively more complex
feature detectors using this reconstruction objective, the weights of a
deep network could be initialized to sensible values.  A final layer of
output units could then be added to the top of the network and the whole
deep system could be fine-tuned using standard
backpropagation\cite{Science}. This worked remarkably well for recognizing
handwritten digits especially when the amount of labeled data was very
limited\cite{Bengio-nips-2006-small}.

The first major application of this pre-training approach was in speech
recognition and it was made possible by the advent of fast Graphics
Processing Units (GPUs) that were convenient to program~\citep{RainaICML09-small}.  
In 2009, the approach was used to map short temporal windows of coefficients extracted
from a soundwave to a set of probabilities for the various fragments of
speech that might be represented by the frame in the center of the window.
It achieved record-breaking results on a standard speech recognition
benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly
developed to give record-breaking results on a large vocabulary
task~\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
developed by many of the major speech groups\cite{SPS} and were already
being deployed in Android phones.  For smaller datasets, unsupervised
pre-training leads to significantly better generalization~\citep{Erhan+al-2010-small}, 
but once deep
learning had been rehabilitated, it turned out that large labeled datasets
did not need the pre-training stage, provided the scales of the initial
random weights were carefully chosen to avoid very large or very small
initial gradients.

Interest in recurrent networks (which are deep in time) was revived by
``Long Short Term Memory'' networks (LSTMs) that use special hidden units
whose natural behaviour is to remember inputs for a long
time~\cite{Hochreiter+Schmidhuber-1997}.  
A linear ``memory cell'' interacts with other units
via gated connections whose weights are multiplied by the state of a
logistic gating unit that has an activation between $0$ and $1$. The input
vector has little effect on the state of a memory cell if the cell's input
gate has a low activation so memory cells can learn to selectively ignore
the input.  Similarly, a memory cell has little effect on the output if its
output gate has low activation. Each memory cell also has a connection to
itself (it is a leaky neuron) at the next time-step that has a weight of 1 so it tends to copy its
own real-valued state, but this self-connection is gated by a ``keep gate'' so
the stored value leaks quickly unless the keep-gate has an activation
near $1$.  The gates themselves have their activities determined by
weighted inputs that they receive from the input vectors and all of the
memory cells.

When LSTMs were first described in
English~\cite{Hochreiter+Schmidhuber-1997} their complexity was off-putting
and, despite their impressive ability to learn very long-range temporal
structure on toy problems, they were largely ignored until they achieved
excellent results in a competition to read cursive
hand-writing~\cite{Graves}. They have subsequently proved to be very
effective for tasks that have sequential input.  They are now better than
deep feed-forward networks for speech recognition~\cite{gravestimit} and
unlike feed-forward nets they can implement an entire speech recognition
system that goes all the way from coefficents derived from the soundwave to
the sequence of characters in the transcription~\cite{gravestimit}.  They
are also what is currently used for the encoder and decoder networks that
perform so well at machine
translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

@@

When deep convolutional networks were applied to a dataset of about a
million images from the web that contained a thousand different classes
they achieved spectacular results, almost halving the error rates of the
best computer vision sytems~\cite{kriz}.  Some of the images and the
network's predictions are shown in figure ??.  The major factors in this
success were the large labeled dataset~\cite{imagenet_cvpr09}, the use of fast GPUs,
and the convolutional early layers.  It had already been shown that deep
supervised nets with rectified linear units could be trained without
unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
These rectified linear units made the
training much faster and the fully connected final layers were prevented
from overfitting the peculiarities of the training set by randomly omitting
hidden units during training~\cite{Srivastava14}. Deep convolutional nets
have quickly replaced the carefully engineered pipelines full of
hand-designed features developed by the computer vision community and they
are now deployed in many computer vision applications, including image
tagging, object detection, face recognition, facial expression recognition
and scene parsing.

@@

\begin{figure}[ht]
%\centerline{\includegraphics[width=0.7\linewidth]{}}
\caption{Example of captions generated by a recurrent network taking
as extra input the representation extracted by a deep convolutional
net from an image, with the whole system trained to ``translate''
images into captions. Reproduced from~\citet{?}.
}
\label{fig:caption-generation}
\end{figure}

@@

Deep learning is now the method of choice for many perceptual tasks and is
making major inroads in natural language processing where a huge amount of
messy commonsense knowledge is required. Rapid progress is also being made
in applying deep learning, in particular BPTT, to relatively pure symbol processing tasks.
An impressive recent example is the ``Neural Turing Machine'' which is a
recurrent neural network that can choose when and what to write to or read
from an external memory~\cite{NTM}. Among other things, this can learn to
ouput a sorted list of symbols when its input consists of an unsorted
sequence in which each symbol is accompanied by a real value that indicates
its priority in the list.  To achieve this it must learn a sorting program
purely from examples of the input and output sequences.  A different system
that also uses BPTT can learn to map the sequence of characters in a simple
program written in Python to the sequence of characters that the program
would print~\cite{Woj}. Rather surprisingly, BPTT can learn about {\it for
  loops} and can also learn to perform addition and multiplication of
strings containing several decimal digits. Achieving this kind of behaviour
seemed ludicrously optimistic when BPTT was first proposed.

@@

This review has focused on supervised learning because of its recent
achievements and we have ignored much interesting work on unsupervised
learning procedures for deep neural networks~\cite{deep-boltz, wake-sleep,
  google-cat, denoising-auto, ranzato-pami, more recent}. In the long run,
however, we believe unsupervised learning will be essential for allowing
deep networks to generalize well by creating internal representations that
untangle the many different factors that interact to produce images, speech
and documents~\cite{Yann,Bengio-Courville-Vincent-TPAMI2013}.

@@

\subsection{Deep unsupervised learning}
Hinton, G., Osindero, S., \& Teh, Y. W. (2006)\\ A fast learning algorithm
for deep belief nets.\\ {\it Neural Computation}, 18(7), 1527-1554.

Erhan, D., Bengio, Y., Courville, A., Manzagol, P. A., Vincent, P., \&
Bengio, S. (2010)\\ Why does unsupervised pre-training help deep
learning?\\ {\it The Journal of Machine Learning Research}, {\bf 11},
625-660.

@@

\subsection{Deep speech recognition}

@@

Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012)\\ Imagenet
classification with deep convolutional neural networks.\\ In Advances in
Neural Information Processing Systems (1097-1105).

@@

Sutskever, I., Vinyals, O., \& Le, Q. V. (2014)\\ Sequence to sequence
learning with neural networks.\\ {\it arXiv preprint}, arXiv:1409.3215.

Bahdanau, D., Cho, K., \& Bengio, Y. (2014)\\ Neural machine translation by
jointly learning to align and translate.  {\it arXiv preprint},
arXiv:1409.0473.

Vinyals, O., Toshev, A., Bengio, S., \& Erhan, D. (2014)\\ Show and Tell: A
Neural Image Caption Generator.\\ {\it arXiv preprint}, arXiv:1411.4555.

>>>>

So far, most applications of deep learning use feed-forward neural
networks, but for tasks that involve sequential inputs it is often better
to use recurrent neural networks. These are given one element from an input % YB term-->element
sequence at each time step and their recurrent internal ``hidden'' state
depends on the whole history of the input sequence. Recurrent networks are
very good at predicting the next character in a string~\citep{Sutskever-et-al-ICML2011} or
the next word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
more complex tasks.  After reading an English sentence one word at a time,
for example, an English``encoder'' network can be trained so that the final
state vector of its hidden neurons is a good representation of the thought
expressed by the sentence.  This thought vector can then be used as the
initial hidden state of a trained French ``decoder'' network which outputs
a probability distribution for the first word of the French translation. If
a particular first word is chosen (e.g., according to that probability)
and provided as input to the decoder
network it will then output a probability distribution for the second word
of the translation and so on until a full stop is 
chosen~\cite{Vinyals-et-al-arxiv2014,Bahdanau-et-al-arxiv2014}. 
%YB added a clarification
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

@@

Patterns can be classified by first converting each raw input vector into a
vector of feature activations and then collecting evidence for each
possible class by computing a weighted sum of the features. If good
features can be designed by hand, it is relatively easy to learn
class-specific weights that optimize
discrimination~\citep{Rosenblatt57}. Widely-used examples of such linear
classification methods are logistic regression and linear support vector
machines~\citep{logisticregression,Boser92}.

However, even with good hand-crafted features, many interesting tasks
cannot be solved by linear classifiers. A common method to get around this
problem is to transform the feature vector into a higher dimensional vector
through a non-linear mapping, exploiting the fact that collections of
vectors are more likely to be easily spearated when they are spread around
a high-dimensional space~\cite{cover}. One clever way to do so is to use a
``kernel function'' that measures the similarity between two vectors.
Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input
vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
optimize the choice of training inputs used to define features and the
class-specific weights associated with the features~\cite{Boser92}. The training
inputs used to define features are called ``support vectors'' and Support
Vector Machines~\citep{Boser92} have been very successful, especially for
tasks with a relatively small number of training examples in which features
are individually informative. They do not, however, perform well for tasks
such as object recognition in which the input space is high-dimensional and
the input-output mapping is so complex that computing similarities with a
simple kernel function is ineffective.

@@

From the earliest days of pattern recognition~\cite{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,Rumelhart86b,schmidreference}.  
Backpropagation is simply an application of
the chain rule, which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights in a layered
network. The objective function typically measures the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it outputs} of a layer can be computed by working
backwards from the error derivatives with respect to the outputs of the
subsequent layer. Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the parameters (or weights) of each layer (see box
1).

@@

Backpropagation provides an efficient way of computing gradients for all of
the weights in a layered network on a single training case, but there are
still many ways to make use of these gradients. The most obvious method is
to average the gradients over all of the cases in the training set and then
to change each weight in proportion to its negative gradient, which is the
direction where a small change will most improve the objective function. 
When the training set contains millions of cases and is highly redundant, however,
this is needlessly slow. It is much faster to compute the average gradient
on a small ``mini-batch'' of randomly chosen training cases and to update
each weight by this stochastic estimate of its full gradient times a
learning rate which decays as learning proceeds. Stochastic gradient
descent can be improved by using the history of the individual weight
gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
or to provide additional increments that depend on earlier gradients~\cite{momentum}
but these techniques are still the subject of controversy and are beyond
the scope of this review.  However, it is now widely accepted that some
form of stochastic gradient descent is the best bet for large, redundant
datasets.

@@

The connectivity of the network is restricted so that each of the one-of-N
input vectors must first be converted to a ``word-vector'' before being
used by subsequent layers to give a high probability to the target next
word.  The network learns word-vectors that contain many active components
each of which can be interpreted as a separate feature of the word. This
was first demonstrated\cite{RHW} using sequences of three words, such as
(Colin, Mother, Victoria), derived from two family trees. The learned word
vector for Colin contained ``semantic'' features representing which family
tree Colin was in, what branch of that tree he was in, and from what generation
he was~\citep{Hinton-et-al-PDP1986}. These semantic features were not explicitly present
in the input.  They were discovered by the learning procedure as a good way
of factorizing the structured relationships between the input and output
symbols into multiple ``micro-rules''.  For example, given Colin's
generation and the learned feature of Mother that the output should be one
generation higher than the input, we can already restrict the answer to
people in a specific generation. The ability of deep learning to discover
many such micro-rules allows it to generalize well to novel combinations of
individually familiar context words.

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or vanish~\citep{Bengio-et-al-TNN1994,Hochreiter91-small}.

@@

The obvious weakness of gradient descent is that it can get trapped in poor
local minima, but in high-dimensional weight-spaces this is much less
common than getting stuck close to saddle points where the error surface
curves upwards in most directions but downwards in a few. Recent
results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-arxiv2014} strongly suggest
that in deep networks almost all of these local minima give very similar
performance so it does not matter which one we find.  This goes a long way
towards explaining why stochastic gradient descent works so well for deep
neural networks.

Interest in deep feed-forward networks was revived in
2005 and 2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small} 
by the introduction
of unsupervised learning procedures that could create layers of feature
detectors without requiring labeled data. The objective in learning each
layer of feature detectors was to be able to reconstruct or model the
activities of feature detectors (or raw inputs) in the layer below.  By
``pre-training'' several layers of progressively more complex feature
detectors using this reconstruction objective, the weights of a deep
network could be initialized to sensible values.  A final layer of output
units could then be added to the top of the network and the whole deep
system could be fine-tuned using standard
backpropagation\cite{Hinton-Science2006}. This worked remarkably well for recognizing
handwritten digits especially when the amount of labeled data was very
limited~\citep{Bengio-nips-2006-small,Erhan+al-2010-small}.

The first major application of this pre-training approach was in speech
recognition and it was made possible by the advent of fast Graphics
Processing Units (GPUs) that were convenient to program~\citep{RainaICML09-small}.  
In 2009, the approach was used to map short temporal windows of coefficients extracted
from a soundwave to a set of probabilities for the various fragments of
speech that might be represented by the frame in the center of the window.
It achieved record-breaking results on a standard speech recognition
benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly
developed to give record-breaking results on a large vocabulary
task~\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
developed by many of the major speech groups\cite{SPS} and were already
being deployed in Android phones.  For smaller datasets, unsupervised
pre-training helps to prevent overfitting~\citep{Erhan+al-2010-small}, 
leading to significantly better generalization when the number of
labeled examples is small. Once deep
learning had been rehabilitated, it turned out that large labeled datasets
did not need the pre-training stage, provided the scales of the initial
random weights were carefully chosen to avoid very large or very small
initial gradients.

Interest in recurrent networks (which are deep in time) was revived by
``Long Short Term Memory'' networks (LSTMs) that use special hidden units
whose natural behaviour is to remember inputs for a long
time~\citep{Hochreiter+Schmidhuber-1997}.  
A linear ``memory cell'' interacts with other units
via gated connections whose weights are multiplied by the state of a
logistic gating unit that has an activation between $0$ and $1$. The input
vector has little effect on the state of a memory cell if the cell's input
gate has a low activation so memory cells can learn to selectively ignore
the input.  Similarly, a memory cell has little effect on the output if its
output gate has low activation. Each memory cell also has a connection to
itself (it is a leaky neuron) at the next time-step that has a weight of 1 so it tends to copy its
own real-valued state, but this self-connection is gated by a ``keep gate'' so
the stored value leaks quickly unless the keep-gate has an activation
near $1$.  The gates themselves have their activities determined by
weighted inputs that they receive from the input vectors and all of the
memory cells.

When LSTMs were first described in
English~\citep{Hochreiter+Schmidhuber-1997} their complexity was off-putting
and, despite their impressive ability to learn very long-range temporal
structure on toy problems, they were largely ignored until they achieved
excellent results in a competition to read cursive
hand-writing~\cite{Graves}. They have subsequently proved to be very
effective for tasks that have sequential input.  They are now better than
deep feed-forward networks for speech recognition~\cite{gravestimit} and
unlike feed-forward nets they can implement an entire speech recognition
system that goes all the way from coefficents derived from the soundwave to
the sequence of characters in the transcription~\cite{gravestimit}.  They
are also what is currently used for the encoder and decoder networks that
perform so well at machine
translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

@@

When deep convolutional networks were applied to a dataset of about a
million images from the web that contained a thousand different classes
they achieved spectacular results, almost halving the error rates of the
best computer vision sytems~\citep{Krizhevsky-2012-small}.  Some of the images and the
network's predictions are shown in figure~\ref{fig:imagenet-outputs}.  
The major factors in this
success were the large labeled dataset~\citep{imagenet_cvpr09}, the use of fast GPUs,
and the convolutional early layers.  It had already been shown that deep
supervised nets with rectified linear units could be trained without
unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
These rectified linear units made the
training much faster and the fully connected final layers were prevented
from overfitting the peculiarities of the training set by randomly omitting
hidden units during training~\citep{Srivastava14}. Deep convolutional nets
have quickly replaced the carefully engineered pipelines full of
hand-designed features developed by the computer vision community and they
are now deployed in many computer vision applications, including image
tagging, object detection, face recognition, facial expression recognition
and scene parsing.

@@

\begin{figure}[ht]
%\centerline{\includegraphics[width=0.7\linewidth]{}}
\caption{Example of captions generated by a recurrent network taking
as extra input the representation extracted by a deep convolutional
net from an image, with the whole system trained to ``translate''
images into captions. Reproduced from~\cite{?}.
}
\label{fig:caption-generation}
\end{figure}

@@

Deep learning is now the method of choice for many perceptual tasks and is
making major inroads in natural language processing where a huge amount of
messy commonsense knowledge is required. Rapid progress is also being made
in applying deep learning, in particular BPTT, to relatively pure symbol processing tasks.
An impressive recent example is the ``Neural Turing Machine'' which is a
recurrent neural network that can choose when and what to write to or read
from an external memory~\citep{Graves-et-al-arxiv2014}. 
Among other things, this can learn to
ouput a sorted list of symbols when its input consists of an unsorted
sequence in which each symbol is accompanied by a real value that indicates
its priority in the list.  To achieve this it must learn a sorting program
purely from examples of the input and output sequences.  A different system
that also uses BPTT can learn to map the sequence of characters in a simple
program written in Python to the sequence of characters that the program
would print~\citep{Zaremba+Sutskever-arxiv2014}. 
Rather surprisingly, BPTT can learn about {\it for
  loops} and can also learn to perform addition and multiplication of
strings containing several decimal digits. Achieving this kind of behaviour
seemed ludicrously optimistic when BPTT was first proposed.

@@

This review has focused on supervised learning because of its recent
achievements and we have ignored much interesting work on unsupervised
learning procedures for deep neural 
networks~\citep{Salakhutdinov2009-small,wake-sleep,
  google-cat, denoising-auto, ranzato-pami, more recent}. In the long run,
however, we believe unsupervised learning will be essential for allowing
deep networks to generalize well by creating internal representations that
untangle the many different factors that interact to produce images, speech
and documents~\cite{Yann,Bengio-Courville-Vincent-TPAMI2013}.

@@

%\subsection{Deep unsupervised learning}
%Hinton, G., Osindero, S., \& Teh, Y. W. (2006)\\ A fast learning algorithm
%for deep belief nets.\\ {\it Neural Computation}, 18(7), 1527-1554.
% --> \citep{Hinton06-small}

%Erhan, D., Bengio, Y., Courville, A., Manzagol, P. A., Vincent, P., \&
%Bengio, S. (2010)\\ Why does unsupervised pre-training help deep
%learning?\\ {\it The Journal of Machine Learning Research}, {\bf 11},
%625-660.
% --> \citep{Erhan+al-2010-small}

@@

%\subsection{Deep speech recognition}

@@

%Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012)\\ Imagenet
%classification with deep convolutional neural networks.\\ In Advances in
%Neural Information Processing Systems (1097-1105).
% --> \cite{Krizhevsky-2012-small}

@@

%Sutskever, I., Vinyals, O., \& Le, Q. V. (2014)\\ Sequence to sequence
%learning with neural networks.\\ {\it arXiv preprint}, arXiv:1409.3215.
% --> \citep{Sutskever-et-al-NIPS2014}

%Bahdanau, D., Cho, K., \& Bengio, Y. (2014)\\ Neural machine translation by
%jointly learning to align and translate.  {\it arXiv preprint},
%arXiv:1409.0473.
% --> \citep{Bahdanau-et-al-arxiv2014}

%Vinyals, O., Toshev, A., Bengio, S., \& Erhan, D. (2014)\\ Show and Tell: A
%Neural Image Caption Generator.\\ {\it arXiv preprint}, arXiv:1411.4555.
% --> \citep{Vinyals-et-al-arxiv2014}
